---
title: "Lab_5 time series and a map"
author: "Yingfei Jiang"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach packages
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewable)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make description all lowercase
- Only keep observations where "description" variable contains "consumption"
- Remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) #! is gonna do the opposite
```

```{r}
renew_date <- renew_clean %>%
  mutate(yr_mo_day = lubridate :: parse_date_time(yyyymm, "ym")) %>%
  mutate(month_sep = yearmonth(yr_mo_day)) %>%
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# Make a version wher I have month and year in separate columns

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at it:

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, y = value, group = description)) +
  geom_line(aes(color = description))

renew_gg
```

updating colors with paletteer palettes:

```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::figmtn")
```

### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our time series data in couple of different ways
```{r}
renew_ts %>%  autoplot(value)
renew_ts %>%  gg_subseries(value)
```

```{r}
renew_ts %>% gg_season(value)

ggplot(data = renew_parsed, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description,
             ncol = 1,
             scale = "free",
             strip.position = "right")
```

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
```

### What if i want the quarterly average consumptions for hydro?
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarise(avg_consumption = mean(value))
```

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season (window = 5))) #Lowest smoothing

components(dcmp) %>%  autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

```{r}
hydro_ts %>% 
  ACF (value) %>% 
  autoplot()

```

# DANGER DANGER

```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ets = ETS(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>%  autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### Make a World Map!

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"), layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

